The truncated KdV framework for modeling anomalous waves induced by abrupt depth changes

### 2019-06-25 WavesData-2019/LatexNotes/notes.tex
NM
N

@@

\subsection{KdV with depth change}
In dimensional form the Korteweg–de Vries equation is given by
\begin{equation}
2 \eta_t + \frac{3 c}{h_0} \eta \eta_{\xi} + \frac{c h_0^2}{3} \eta_{\xi \xi \xi} = 0
\end{equation}

@@

The dimensionless KdV equation is
\begin{align}
\label{varKdV}
u_t + \ampp D^{-3/2} \, u u_x + \mu D^{1/2} \, u_{xxx} = 0
\end{align}
with the dimensionless parameters
\begin{align}
\label{ndKdV}
\ampp = \eps \delta^{-2} = \frac{ag}{h_0^2 f_p^2} , \qquad
\mu = \frac{4 \pi^2}{9 \lamfac^2}
\end{align}
Here, the displacement has been scaled on the characteristic wave amplitude, $u = \eta/a$, where $a = \etastd= \mean{\eta^2}^{1/2}$.
As a consequence, $u$ has fixed energy
\begin{equation}
E[u] = \frac{1}{2} \intt u^2 \dx = \frac{1}{2}
\end{equation}

@@

@@

The statistical matching condition is
\begin{equation}
\meanm{\Hp} = \meanp{\Hp}
\end{equation}
An experimental observation is that the incoming skewness is small, thus
\begin{equation}
\meanm{H_3} \approx 0
\end{equation}
which immediately gives
\begin{equation}
\ampp D^{-3/2} \meanp{H_3} = \mu D^{1/2} \left( \meanp{H_2} - \meanm{H_2} \right)
\end{equation}
or equivalently
\begin{equation}
\meanp{H_3} = \frac{\mu D^{2}}{\ampp} \, \Delta \mean{H_2}
\end{equation}
where $\Delta \mean{H_2} =  \meanp{H_2} - \meanm{H_2}$  indicates the difference between the upstream and downstream values.

To convert to dimensional, i.e.~experimental, values we use
\begin{align}
& \intt u^3 \dx = \frac{\mean{\eta^3}}{\mean{\eta^2}^{3/2}} = \skw(\eta) \\
& \intt u_x^2 \dx = \left(\frac{\lambda_0}{a} \right)^2 \mean{\eta_x^2} 
= \left(\frac{\lambda_0}{a} \right)^2 \var(\eta_x)
\end{align}
Then we have the relationship
\begin{equation}
\frac{\skw(\eta)}{\Delta \var(\eta_x)} =
3 \mu D^2 \left( \frac{h_0}{a} \right)^3 = 3 \eps^{-3} \mu D^2
\end{equation}
In particular, the ratio on the left, which can be measured in the experiments, is predicted to scale as the inverse cube of the wave amplitude (as well as the square of the depth ratio).

>>>>

% New
\newcommand{\depth}{d}
\newcommand{\dpm}{\depth^{-}}
\newcommand{\dpp}{\depth^{+}}

@@

\subsection{KdV with depth change}
Consider waves propagating unidirectionally in shallow water. Consider the surface displacement $\eta(x,t)$ and the reference frame moving with the local wave speed $\xi = x - ct$, where $c = \sqrt{g \depth}$ is the wave speed, $g$ gravity, and $\depth$ the local depth.
To first-correction in small amplitude, surface displacements are governed by the Korteweg–de Vries equation, which in dimensional form is given by
\begin{equation}
2 \eta_t + \frac{3 c}{\depth} \eta \eta_{\xi} + \frac{c \depth^2}{3} \eta_{\xi \xi \xi} = 0
\end{equation}

We will primarily consider the case in which waves originate from a region of constant depth, encounter an abrupt depth change, and then continue in another region of constant depth. Thus, depth will be piecewise constant
\begin{align}
\depth = 
\begin{cases}
\it \dpm \quad \mbox{if } x<0 \\
\it \dpp \quad \mbox{if } x>0
\end{cases}
\end{align}
Most often, we will consider waves moving into shallower depth, so that $\dpp < \dpm$.

@@

We introduce the following characteristic scales
\begin{align}
\mbox{\em characteristic wave amplitude} \qquad
a = \etastd = \mean{\eta^2}^{1/2} \\
\mbox{\em characteristic wave length} \qquad
 = \etastd = \mean{\eta^2}^{1/2}
\end{align}

\vsp{10}
The dimensionless KdV equation is
\begin{align}
\label{varKdV}
u_t + \ampp D^{-3/2} \, u u_x + \mu D^{1/2} \, u_{xxx} = 0
\end{align}
with the dimensionless parameters
\begin{align}
\label{ndKdV}
\ampp = \eps \delta^{-2} = \frac{ag}{h_0^2 f_p^2} , \qquad
\mu = \frac{4 \pi^2}{9 \lamfac^2}
\end{align}
Here, the displacement has been scaled on the characteristic wave amplitude, $u = \eta/a$, where $a = \etastd= \mean{\eta^2}^{1/2}$.
As a consequence, $u$ has fixed energy
\begin{equation}
E[u] = \frac{1}{2} \intt u^2 \dx = \frac{1}{2}
\end{equation}

@@

NOTE: We need to include the bandwidth, 2 Hz, in the theory somehow. For example, the peak forcing frequency and bandwidth together can be used to estimate the decay in the spectrum.

@@

The statistical matching condition is
\begin{equation}
\meanm{\Hp} = \meanp{\Hp}
\end{equation}
An experimental observation is that the incoming skewness is small, thus
\begin{equation}
\meanm{H_3} \approx 0
\end{equation}
which immediately gives
\begin{equation}
\ampp D^{-3/2} \meanp{H_3} = \mu D^{1/2} \left( \meanp{H_2} - \meanm{H_2} \right)
\end{equation}
or equivalently
\begin{equation}
\label{H3H2}
\meanp{H_3} = \frac{\mu D^{2}}{\ampp} \, \Delta \mean{H_2}
\end{equation}
where $\Delta \mean{H_2} =  \meanp{H_2} - \meanm{H_2}$  indicates the difference between the upstream and downstream values.

To convert to dimensional, i.e.~experimental, values we use
\begin{align}
& \intt u^3 \dx = \frac{\mean{\eta^3}}{a^3} = 
\frac{\mean{\eta^3}}{\mean{\eta^2}^{3/2}} = \skw(\eta) \\
& \intt u_x^2 \dx = \left(\frac{\lambda_0}{a} \right)^2 \mean{\eta_x^2} 
= \left(\frac{\lambda_0}{a} \right)^2 \var(\eta_x)
\end{align}
Then \eqref{H3H2} gives the relationship
\begin{equation}
\frac{\skw(\eta)}{\Delta \var(\eta_x)} =
3 \mu D^2 \left( \frac{h_0}{a} \right)^3 = 3 \eps^{-3} \mu D^2
\end{equation}
In particular, the ratio on the left, which can be measured in the experiments, is predicted to scale as the inverse cube of the wave amplitude (as well as the square of the depth ratio).
